---
  title: "An example about how to evaluate manually constructed models with the autohrf package"
  author: "Jure Demšar, Nina Purg and Grega Repovš"
  date: "`r Sys.Date()`"
  output:
    html_vignette:
    toc: yes
---

<!--
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{b_reaction_time: Bayesian reaction time model}
-->

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# knitr options
knitr::opts_chunk$set(fig.width=6, fig.height=4.5)
options(width=800)
```

The **autohrf** package is not only useful for preparing model specifications and then automatically finding the models that best fit the underlying data. In this example we will show how you can use the **autohrf** package to investigate the quality of manually constructed models.

Let us start this example by loading required libraries and the data from the spatial working memory experiment (SWM).

```{r}
# libs
library(autohrf)

# load the data
df <- swm
head(df)
```

The loaded data frame has 504 observations, each with 3 variables (roi, t, and y) **roi** denotes the region of interest, **t** the time stamp and **y** the value of the BOLD signal. Note that input data for the **autohrf** package should be always organized in this manner.

Next, we will construct three different models, one with 3 events, one with 4 events and one with 5 events. When manually constructing events we need to create a data frame which has an entry (observation) for each of the events in the model. For each of the events we need to provide its name, its start time and its duration.

```{r}
# 3 events: encoding, delay, response
model3 <- data.frame(event      = c("encoding", "delay", "response"),
                     start_time = c(0,           2.65,    12.5),
                     duration   = c(2.65,        9.85,    3))

# 4 events: fixation, target, delay, response
model4 <- data.frame(event      = c("fixation", "target", "delay", "response"),
                     start_time = c(0,           2.5,      2.65,    12.5),
                     duration   = c(2.5,         0.1,      9.85,    3))

# 5 events: fixation, target, delay, probe, response
model5 <- data.frame(event      = c("fixation", "target", "delay", "probe", "response"),
                     start_time = c(0,           2.5,      2.65,    12.5,    13),
                     duration   = c(2.5,         0.1,      9.85,    0.5,     2.5))
```

Once we construct our models we can use the **convolve_events** function to convolve the HRF signal based on the model with the actual data. Once we have that, we can use the **run_model** function to evaluate which model fits our data best. We can also use the **plot_model** function to visually inspect how model fits the underlying data.

```{r}
# 3 events
em3 <- evaluate_model(df, model3, tr = 2.5)
plot_model(em3)
```

```{r}
# 4 events
em4 <- evaluate_model(df, model4, tr = 2.5)
plot_model(em4)
```

```{r}
em5 <- evaluate_model(df, model5, tr = 2.5)
plot_model(em5)
```

We can see that the best R2 score was achieved by our model with 5 events. So it would make sense to use that one for further assumed GLM modelling.
